Development of a educational and sustainable robotic platform for children and adults

### 2014-04-26 doc/paper.tex
zeugs

\subsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever has an integrated Schmitt trigger and therefore provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

@@

\subsection{IR Proximity Sensor}
Infrared proximity sensors can be used in various ways. When measuring distances using proximity sensors one has to choose the right sensor type since the range varies from a few centimeters to a few meters. Another use of IR proximity sensors is to measure brightness. This usually works with sensors designed for very small distances. When the IR light is emitted it reflects better on bright surfaces than on dark surfaces which do not reflect light. A photodiode for infrared wavelength changes its electrical behaviour on the influence of the light reflected which can be measured using the microcontrollers analog-to-digital converter.

@@

\subsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29m/s at sea level).

@@

\subsection{Accelerometer}
Accelerometers can sense acceleration on one or more axis. This can be used to determine the speed of the robot or the direction in which the robot accelerates (which can be used to determine whether the robot falls/is lifted up).

@@

The MPU-6050 can be interfaced using various protocols, but I2C is the most common way to do so and requires two input/output pins only using software-based I2C or two I2C hardware pins. Libraries with permissive license exist which ease the use of the accelerometer. For advanced usage scenarios a integrated DMP (Digitial Motion Processor) can process raw values and interface additional sensors like a compass.
\subsection{Gyroscope}
Gyroscopes are used to determine the orientation. Applications for gyroscopes are balancing and position calculation (usually in combination with an accelerometer). Integrated circuits such as the MPU-6050 mentioned before usually combine a gyroscope and an accelerometer into one single device.

@@

As mentioned in the previous section the MPU-6050 is easily interfaced using various protocols of which I2C is the common way.
\subsection{Compass}
A compass allows the robot to sense its orientation relative to the magnetic field of the earth. Technically speaking a integrated circuit providing measurement of the magnetic field of the earth are a special form of magnetometers.

@@

Advantages over other battery types exist in energy density and the high cell voltage of 3.7 Volt which reduces the amount of cells needed. The high current during discharge is well suited for driving variable current sources such as DC motors. Its small size and low cost of about 10 USD for a 3.7 Volt 1000mAh cell with a weight of less than 10 gramm is easy to handle.
\subsubsection{Conclusion}
Even though lithium-polymer batteries expose a greater risk when handled wrong (overcharging, over-discharging and over-discharge-current), they provide higher rated discharge current and a high cell voltage. The size of lithium-polymer batteries is compact and security measures  exist which prevent harm up to a certain extend.
\section{Robot Construction}
The design and construction process of the robot is based on the evaluation and evaluation critera. The design process is centered on critera set by the components used and the previous evaluation.

@@

The robot constructed will feature two microcontroller where one acts as the master and the other as the slave device. The master is fully compatible to the Arduino Leonardo and provides Arduino headers for shields. In addition an extended header is included to match the standard 2.54 inch prototyping board. Communication between the microcontroller is realized using I2C. The second microcontroller controls sensors, movement and optional components such as networking. The schematics are in the appendix.

@@

@@

A typical Gripper is moved

\subsection{PCB design}

@@

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{5}    & With a price tag of 27.68\euro for the base kit and 37.91\euro for the kit with all extensions\footnote{without VAT, based on mass production calculation for hundred units.} the robots costs are below the maximum price of 100\euro.\\
Educational     & \stars{4}     & Assembling the robot and its components includes a unique learning experience and allows the explanation of each part. The Arduino compatibility enables the usage of comprehensive tutorials and how-to documentation. Learning material for teachers is not yet supplied. \\
Sustainable       & \stars{5}     &  The robot integrates charging circuit for the Lithium-polymer batteries which eliminates the need of a dedicated battery charger and therefore reduces the amount of electrical waste. All components can be easily replaced by beginners by either changing the pre-soldered modules or solder a new component. No tantalum capacitors are included.\\
Extendable & \stars{5}      &  Sensors and a display can be directly plugged in to extend the functionality of the robot. Unused ports are broken out onto pin headers. Fully compatible Arduino headers allow the usage of Arduino shields.\\
Trouble-free & \stars{4} & Building the robot is made easy doe to super sized SMD components. When possible 1206 sized parts are used. Some ICs on the board are hard to solder (e.g. the MCP 1650) and should come pre-assembled to avoid frustration.\\
Open-Source & \stars{5} & The design file for laser cutting, the pcb schematic and board files are available under a Creative Commons Attribution Share-Alike license.\\
\bottomrule
\caption{Robot evaluation}
\label{tbl:arduinorobot_eval}
\end{longtable}

@@

\begin{longtable}{p{0.5\textwidth}p{0.15\textwidth}p{0.2\textwidth}p{0.1\textwidth}}
\toprule
Part & Amount & Cost per part\footnote{Average price when ordered in quantity of 100.} & Sum \\
\midrule
\multicolumn{4}{c}{\textbf{US Distributor}\footnote{Supplied via \url{http://digikey.de}, accessed on 16th of March 2014. Prices are based on quantities of 100 units.}}\\
\\
CAP 18pF 1206 50V & 4 & 0.04\euro & 0.16\euro \\ 
CAP 0.1uF 1206 50V & 8 & 0.04\euro & 0.32\euro \\
CAP 1uF 1206 50V & 5 & 0.05\euro & 0.25\euro \\
CAP 4.7uF 1206 16V & 2 & 0.06\euro & 0.12\euro \\
CAP 10uF 1206 16V & 3 & 0.05\euro & 0.15\euro \\
CAP 47uF 1206 10V & 2 & 0.28\euro & 0.56\euro \\
RES 22 OHM 1206 1\% & 4 &  0.01\euro & 0.04\euro \\
RES 100 OHM 1206 5\% & 5 & 0.01\euro & 0.05\euro \\
RES 470 OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 1K OHM 1206 5\% & 7 & 0.01\euro & 0.07\euro \\
RES 2K OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 3.09K OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 10K OHM 1206 5\% & 8 & 0.01\euro & 0.08\euro \\
B330A & 1 & 0.25\euro & 0.25\euro \\
LED 1206 & 7 & 0.11\euro & 0.77\euro \\
PTC RESET 15V .5A SMD 1206 & 2 & 0.26\euro & 0.52\euro \\
MOSFET N-CH 20V 4.2A SOT-23 & 2 & 0.19\euro & 0.38\euro \\
FERRITE 3A 100 OHM 1206 SMD & 1 & 0.05\euro & 0.05\euro \\
INDUCTOR 3.3UH 1210 SMD & 1 & 0.15\euro & 0.15\euro \\
CRYSTAL 16MHZ 20PF SMD & 2 & 0.25\euro & 0.50\euro \\
ATMEGA32U4 44TQFP & 2 & 2.80\euro & 5.60\euro \\
LDO 3.3V SOT 23 & 1 & 0.33\euro & 0.33\euro \\ 
MCP1650 & 1 & 0.79\euro & 0.79\euro \\
MCP73831 & 1 & 0.33\euro & 0.33\euro \\
74HC595D SO16 & 2 & 0.20\euro & 0.40\euro \\
74HC4051 SO16 & 1 & 0.21\euro & 0.21\euro \\
ULN2803 SO18 & 1 & 0.54\euro & 0.54\euro \\
\\
\multicolumn{4}{c}{\textbf{Chinese Distributor}\footnote{\label{edragon}Supplied via Chinese distributors found on \url{http://alibaba.com} on the 16th of March 2014. The price is based on quantities of one, dollar prices were converted to Euro with 1\euro = 1.3USD }}\\
\\
CONN 02-JST-XH & 1 & 0.02\euro & 0.02\euro \\
CONN 05-JST-XH & 2 & 0.05\euro & 0.10\euro \\
CONN 2.54 FEMALE 1x06 & 2 & 0.07\euro & 0.14\euro \\
CONN 2.54 FEMALE 1x08 & 1 & 0.08\euro & 0.08\euro \\
CONN 2.54 FEMALE 1x10 & 2 & 0.10\euro & 0.20\euro \\
CONN 2.54 MALE 1x03 & 3 & 0.01\euro & 0.03\euro \\
CONN 2.54 MALE 1x04 & 1 & 0.01\euro & 0.01\euro \\
CONN 2.54 MALE 2x03 & 4 & 0.01\euro & 0.04\euro \\
CONN USB MINI B & 2 & 0.04\euro & 0.08\euro \\
TACTILE SWITCH & 2 & 0.02\euro & 0.04\euro \\
TOGGLE SWITCH & 1 & 0.03\euro & 0.03\euro \\
MICRO SERVO & 2 & 2.04\euro & 4.08\euro \\
IR RECV 38KHZ & 1 & 0.10\euro & 0.10\euro \\
STEPPER MOTOR 28-BYJ & 2 & 1.39\euro & 2.78\euro \\
HC-SR04 ULTRASONIC & 1 & 1.38\euro & 1.38\euro \\
1100mAH LIPO BATTERY & 1 & 1.84\euro & 1.84\euro \\
\\
\multicolumn{4}{c}{\textbf{German Distributor}\footnote{Supplied via German distributors or made in Germany.}}\\
\\
LASERCUT PARTS\footnote{Price is based on the raw material price of 0.60\euro per DIN A3 sheet of 4mm plywood. Lasercutting costs are not included. } & 1 & 0.60\euro & 0.60\euro \\
PCB\footnote{Price based on calculation at \url{http://multi-circuit-boards.eu} with the PCB dimensions, double-sided with solder stop.} & 1 & 3.66\euro & 3.66\euro \\ \hline\hline
Total cost& & & 27.68\euro \\
\bottomrule
\caption{Bill of Material for required parts (assuming mass production of 100 units)}
\label{tbl:bom}
\end{longtable}

\begin{longtable}{p{0.5\textwidth}p{0.15\textwidth}p{0.2\textwidth}p{0.1\textwidth}}
\toprule
Part & Amount & Cost per part\footnote{Items supplied via Chinese distributors found on \url{http://alibaba.com} on the 16th of March 2014. The price is based on quantities of one, dollar prices were converted to Euro with 1\euro = 1.3USD } & Sum \\
\midrule
DISPLAY SPI 1.8 INCH & 1 & 4.47\euro & 4.47\euro \\
NRF24L01+ & 1 & 0.92\euro & 0.92\euro \\
MPU6050 GYRO + ACCEL & 1 & 2.76\euro & 2.76\euro \\
HMC-5883L & 1 & 1.84\euro & 1.84\euro \\
CONN 2.54 FEMALE 1x04 & 1 & 0.04\euro & 0.04\euro \\
CONN 2.54 FEMALE 1x05 & 1 & 0.05\euro & 0.05\euro \\
CONN 2.54 FEMALE 1x08 & 2 & 0.05\euro & 0.10\euro \\
CONN 2.54 FEMALE 2x04 & 1 & 0.05\euro & 0.05\euro \\
\hline \\
Total cost& & & 10.23\euro \\
\bottomrule
\caption{Bill of Material for optional parts (assuming mass production of 100 units)}
\label{tbl:bom}
\end{longtable}
\includepdf[pages={1}]{robot.pdf}

>>>>

\subsubsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever has an integrated Schmitt trigger and therefore provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

@@

\subsubsection{IR Proximity Sensor}
Infrared proximity sensors can be used in various ways. When measuring distances using proximity sensors one has to choose the right sensor type since the range varies from a few centimeters to a few meters. Another use of IR proximity sensors is to measure brightness. This usually works with sensors designed for very small distances. When the IR light is emitted it reflects better on bright surfaces than on dark surfaces which do not reflect light. A photodiode for infrared wavelength changes its electrical behaviour on the influence of the light reflected which can be measured using the microcontrollers analog-to-digital converter.

@@

\subsubsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29m/s at sea level).

@@

\subsubsection{Accelerometer}
Accelerometers can sense acceleration on one or more axis. This can be used to determine the speed of the robot or the direction in which the robot accelerates (which can be used to determine whether the robot falls/is lifted up).

@@

The MPU-6050 can be interfaced using various protocols, but I2C is the most common way to do so and requires two input/output pins only using software-based I2C or two I2C hardware pins. Libraries with permissive license exist which ease the use of the accelerometer. For advanced usage scenarios a integrated DMP (Digitial Motion Processor) can process raw values and interface additional sensors like a compass.
\subsubsection{Gyroscope}
Gyroscopes are used to determine the orientation. Applications for gyroscopes are balancing and position calculation (usually in combination with an accelerometer). Integrated circuits such as the MPU-6050 mentioned before usually combine a gyroscope and an accelerometer into one single device.

@@

As mentioned in the previous section the MPU-6050 is easily interfaced using various protocols of which I2C is the common way.
\subsubsection{Compass}
A compass allows the robot to sense its orientation relative to the magnetic field of the earth. Technically speaking a integrated circuit providing measurement of the magnetic field of the earth are a special form of magnetometers.

@@

Advantages over other battery types exist in energy density and the high cell voltage of 3.7 Volt which reduces the amount of cells needed. The high current during discharge is well suited for driving variable current sources such as DC motors. Its small size and low cost of about 10 USD for a 3.7 Volt 1000mAh cell with a weight of less than 10 gramm is easy to handle.
\subsubsection{Conclusion}
Even though lithium-polymer batteries expose a greater risk when handled wrong (overcharging, over-discharging and over-discharge-current), they provide higher rated discharge current and a high cell voltage. The size of lithium-polymer batteries is compact and security measures  exist which prevent harm up to a certain extend.
\section{Robot Construction}
The design and construction process of the robot is based on the evaluation and evaluation critera. Having the results of the previous evaluation, the design process can focus on the best suited parts.

@@

The robot constructed will feature two microcontroller where one acts as the master and the other as the slave device. The master is fully compatible to the Arduino Leonardo and provides Arduino headers for shields. In addition an extended header is included to match the standard 2.54 inch prototyping board. Communication between the microcontroller is realized using I2C. The second microcontroller controls sensors, movement and optional components such as networking. The schematics are in the appendix. A

@@

Power requirement for the display are 30mA at 3.3V for the backlight and 1mA at 5V for the processing unit according to the datasheet\footnote{\url{http://www.adafruit.com/datasheets/JD-T1800.pdf}}. Having a power subsystem which has spare power of 400mA at 5V and 50mA at 3.3V the display can be safely used.

@@

A typical Gripper is moved using a servo motor and contrarotating gears attached to it. The contrarotating gear enables the gripper to perform a grip and release action. The typical range of 180 degree for a servo motor, represents the typical range needed for grippers to work.

A well suited gripper is the Robotic Claw by Maker named Kepler\footnote{see \url{http://www.thingiverse.com/thing:18339/}}. The 3D printed components are simple, robust and relatively cheap to produce. In combination with the Servo motor the costs are 2.04 Euro for the servo and about 0.20 Euro for the 3D printer filament.

% TODO insert image here

\subsection{PCB design} 
A PCB (short for printed circuit board) is a non-conductive base material with conductive traces connecting electrical components. Electrical components like resistors, microcontrollers or other integrated circuits are soldered to the traces. Components are available in different types, where THT (Through-Hole-Technology) components have the leads soldered to the opposite site through a hole and SMT (Surface-Mount-Technology) components have the leads soldered at the same side. The following illustration shows the difference between both technologies:

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.7\textwidth]{images/part_types_thtsmt.jpg}
  \caption{Illustration of THT (top) and SMT (bottom) parts on a PCB (green) with Solder (grey) (own illustration)}
\end{figure}

PCBs can have multiple layers with copper traces which are connected using vias. A via is hole made conductive using the process of electroplating. If a PCB contains more than one layer, it is likely that vias are used. Leads of THT components can be used as vias as well, since the hole drilled for the lead is usually electroplated as well.

Before a PCB can be created, a schematic must be drawn. A schematic defines the connections between each pin and therefore gives an overview where electrical connections must be made. The schematic for the robot is found in the appendix. It was made using Cadsoft Eagle - a cross platform CAD program for schematic and PCB layout.

When creating the PCB from the schematic, the schematic must be checked for errors first. EAGLE has a Electrical Rule Check (ERC) built in, which performs several checks such as checks for whether each pin has a connection or checks for power pin connections to power lines. In the board view of Cadsoft EAGLE where one can design the PCB, all components are ordered by their appearance beneath the circuit board first.

The main task is to order the components on the PCB, route traces between pins and keep some basic design rules/guidelines:

\begin{itemize}
\item The width of a trace determines its resistance. The thinner the trace the higher is the resistance and therefore it can carry less current or will heat up. The minimum trace width for a manufacturer is usually 6mil (1 mil = 0.001 inch), homemade etched PCBs are limited to a minimum width of about 10-12mil.
\item The polarity of components is important. For ordinary parts include the symbol for pin one and for diodes and polarized capacitors the plus sign on the silkscreen or the copper layer. This eases manual assembly and prevents errors.
\item Traces and pads (place where a lead connects to) should not be placed to close to each other. Soldering is difficult on close traces and pads. 
\item Sharp edges (angle bigger than 45 degree) are considered bad style. If the angle looks like a "v", problems during the etching process may occur. 
\item A ground plane (covering with whole pcb with a ground layer) is good to prevent signalling errors. 
\item Components on screen may look bigger than they are in reality. Adding extra space between components makes them easier to solder. The solder tip should not touch other components.
\item Orientation of components should be equal among equal parts to prevent errors while soldering. This especially holds for polarized parts.
\end{itemize}

Based on these guidelines the PCB is created. At first the components of the Arduino Leonardo compatible unit (first CPU/user CPU) are layouted. Special care is taken with the spacing between the Arduino shield compatible pin headers.\footnote{See the simplified board layout in the Appendix for illustration.}

@@

%todo

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{5}    & With a price tag of 31.28\euro for the base kit and 41.51\euro for the kit with all extensions\footnote{without VAT, based on mass production calculation for hundred units.} the robots costs are below the maximum price of 100\euro.\\
Educational     & \stars{4}     & Assembling the robot and its components includes a unique learning experience and allows the explanation of each part. The Arduino compatibility enables the usage of comprehensive tutorials and how-to documentation. Learning material for teachers is not yet supplied. \\
Sustainable       & \stars{5}     &  The robot integrates charging circuit for the Lithium-polymer batteries which eliminates the need of a dedicated battery charger and therefore reduces the amount of electrical waste. All components can be easily replaced by beginners by either changing the pre-soldered modules or solder a new component. No tantalum capacitors are included.\\
Extendable & \stars{5}      &  Sensors and a display can be directly plugged in to extend the functionality of the robot. Unused ports are broken out onto pin headers. Fully compatible Arduino headers allow the usage of Arduino shields.\\
Trouble-free & \stars{4} & Building the robot is made easy doe to super sized SMD components. When possible 1206 sized parts are used. Some ICs on the board are hard to solder (e.g. the MCP 1650) and should come pre-assembled to avoid frustration.\\
Open-Source & \stars{5} & The design file for laser cutting, the pcb schematic and board files are available under a Creative Commons Attribution Share-Alike license.\\
\bottomrule
\caption{Robot evaluation}
\label{tbl:arduinorobot_eval}
\end{longtable}

@@

\begin{longtable}{p{0.06\textwidth}p{0.43\textwidth}p{0.2\textwidth}p{0.2\textwidth}}
\toprule
Pin&Atmel Pin Name&User CPU&Brain CPU\\
\midrule
1&PE6 (INT.6/AIN0)&Digital pin 7&NRF24L01 IRQ\\
2&UVcc&5V&5V\\
3&D-&USB-&USB2-\\
4&D+&USB+&USB2+\\
5&UGnd&GND&GND\\
6&UCap&UCAP&UCAP2\\
7&VUSB&5V&5V\\
8&(SS/PCINT0) PB0&RXLED&RXLED2\\
9&(PCINT1/SCLK) PB1&SCK&SCK2\\
10&(PDI/PCINT2/MOSI) PB2&MOSI&MOSI2\\
11&(PDO/PCINT3/MISO) PB3&MISO&MISO2\\
12&(PCINT7/OCA0/OC1C/\#RTS) PB7&Digital pin 11 (PWM)&Servo Gripper Rotation\\
13&RESET&RESET&RESET2\\
14&Vcc&5V&5V\\
15&GND&GND&GND\\
16&XTAL2&XTAL2&XTAL22\\
17&XTAL1&XTAL1&XTAL12\\
18&(OC0B/SCL/INT0) PD0&SCL&SCL\\
19&(SDA/INT1) PD1&SDA&SDA\\
20&(RX D1/AIN1/INT2) PD2&Digital pin 0 (RX)&RX2\\
21&(TXD1/INT3) PD3&Digital pin 1 (TX)&TX2\\
22&(XCK1/\#CTS) PD5&TXLED&TXLED2\\
23&GND1&GND&GND\\
24&AVCC&AVCC&AVCC\\
25&(ICP1/ADC8) PD4&Digital pin 4&IR\\
26&(T1/\#OC4D/ADC9) PD6&Digital pin 12&Shift Register RCK\\
27&(T0/OC4D/ADC10) PD7&Digital Pin 6 (PWM)&Servo Gripper Grip\\
28&(ADC11/PCINT4) PB4&Digital pin 8&SD Card CS\\
29&(PCINT5/OC1A/\#OC4B/ADC12) PB5&Digital Pin 9 (PWM)&NRF24L01 CE\\
30&(PCINT6/OC1B/OC4B/ADC13) PB6&Digital Pin 10 (PWM)&NRF24L01 CSN\\
31&(OC3A/\#0C4A) PC6&Digital Pin 5 (PWM)&Servo Ultrasonic Rotation\\
32&(ICP3/CLK0/)C4A) PC7&Digital Pin 13 (PWM) + LED&LED\\
33&(\#HWB) PE2&HWB&HWB\\
34&Vcc1&5V&5V\\
35&GND2&GND&GND\\
36&(ADC7/TDI) PF7&Analog In 0&Ultrasonic Trigger\\
37&(ADC6/TDO) PF6&Analog In 1&Ultrasonic Echo\\
38&(ADC5/TMS) PF5&Analog In 2&TFT DC\\
39&(ADC4/TCK) PF4&Analog In 3&TFT RST\\
40&(ADC1) PF1&Analog In 4&TFT CS\\
41&(ADC0) PF0&Analog In 5&MUX Out\\
42&AREF&AREF&AREF\\
43&GND3&GND&GND\\
44&AVCC1&AVCC&AVCC\\
\bottomrule
\caption{CPU pin mapping for constructed robot}

\end{longtable}

\begin{longtable}{p{0.5\textwidth}p{0.15\textwidth}p{0.2\textwidth}p{0.1\textwidth}}
\toprule
Part & Amount & Cost per part\footnote{Average price when ordered in quantity of 100.} & Sum \\
\midrule
\multicolumn{4}{c}{\textbf{US Distributor}\footnote{Supplied via \url{http://digikey.de}, accessed on 16th of March 2014. Prices are based on quantities of 100 units.}}\\
\\
CAP 18pF 1206 50V & 4 & 0.04\euro & 0.16\euro \\ 
CAP 0.1uF 1206 50V & 8 & 0.04\euro & 0.32\euro \\
CAP 1uF 1206 50V & 5 & 0.05\euro & 0.25\euro \\
CAP 4.7uF 1206 16V & 2 & 0.06\euro & 0.12\euro \\
CAP 10uF 1206 16V & 3 & 0.05\euro & 0.15\euro \\
CAP 47uF 1206 10V & 2 & 0.28\euro & 0.56\euro \\
RES 22 OHM 1206 1\% & 4 &  0.01\euro & 0.04\euro \\
RES 100 OHM 1206 5\% & 5 & 0.01\euro & 0.05\euro \\
RES 470 OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 1K OHM 1206 5\% & 7 & 0.01\euro & 0.07\euro \\
RES 2K OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 3.09K OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 10K OHM 1206 5\% & 8 & 0.01\euro & 0.08\euro \\
B330A & 1 & 0.25\euro & 0.25\euro \\
LED 1206 & 7 & 0.11\euro & 0.77\euro \\
PTC RESET 15V .5A SMD 1206 & 2 & 0.26\euro & 0.52\euro \\
MOSFET N-CH 20V 4.2A SOT-23 & 2 & 0.19\euro & 0.38\euro \\
FERRITE 3A 100 OHM 1206 SMD & 1 & 0.05\euro & 0.05\euro \\
INDUCTOR 3.3UH 1210 SMD & 1 & 0.15\euro & 0.15\euro \\
CRYSTAL 16MHZ 20PF SMD & 2 & 0.25\euro & 0.50\euro \\
ATMEGA32U4 44TQFP & 2 & 2.80\euro & 5.60\euro \\
LDO 3.3V SOT 23 & 1 & 0.33\euro & 0.33\euro \\ 
MCP1650 & 1 & 0.79\euro & 0.79\euro \\
MCP73831 & 1 & 0.33\euro & 0.33\euro \\
74HC595D SO16 & 2 & 0.20\euro & 0.40\euro \\
74HC4051 SO16 & 1 & 0.21\euro & 0.21\euro \\
ULN2803 SO18 & 1 & 0.54\euro & 0.54\euro \\
\\
\multicolumn{4}{c}{\textbf{Chinese Distributor}\footnote{\label{edragon}Supplied via Chinese distributors found on \url{http://alibaba.com} on the 16th of March 2014. The price is based on quantities of one, dollar prices were converted to Euro with 1\euro = 1.3USD }}\\
\\
CONN 02-JST-XH & 1 & 0.02\euro & 0.02\euro \\
CONN 05-JST-XH & 2 & 0.05\euro & 0.10\euro \\
CONN 2.54 FEMALE 1x06 & 2 & 0.07\euro & 0.14\euro \\
CONN 2.54 FEMALE 1x08 & 1 & 0.08\euro & 0.08\euro \\
CONN 2.54 FEMALE 1x10 & 2 & 0.10\euro & 0.20\euro \\
CONN 2.54 MALE 1x03 & 3 & 0.01\euro & 0.03\euro \\
CONN 2.54 MALE 1x04 & 1 & 0.01\euro & 0.01\euro \\
CONN 2.54 MALE 2x03 & 4 & 0.01\euro & 0.04\euro \\
CONN USB MINI B & 2 & 0.04\euro & 0.08\euro \\
TACTILE SWITCH & 2 & 0.02\euro & 0.04\euro \\
TOGGLE SWITCH & 1 & 0.03\euro & 0.03\euro \\
MICRO SERVO & 2 & 2.04\euro & 4.08\euro \\
IR RECV 38KHZ & 1 & 0.10\euro & 0.10\euro \\
STEPPER MOTOR 28-BYJ & 2 & 1.39\euro & 2.78\euro \\
HC-SR04 ULTRASONIC & 1 & 1.38\euro & 1.38\euro \\
1100mAH LIPO BATTERY & 1 & 1.84\euro & 1.84\euro \\
\\
\multicolumn{4}{c}{\textbf{German Distributor}\footnote{Supplied via German distributors or made in Germany.}}\\
\\
LASERCUT PARTS\footnote{Price is based on the raw material (320x230mm) price of 40\euro per square meter of 3mm acrylic. Lasercutting costs are not included. } & 1 & 3.00\euro & 3.00\euro \\
M4 Screw 25mm & 8 & 0.03\euro & 0.24\euro \\
M4 Screw 10mm & 22 & 0.03\euro &  0.66\euro \\
M4 Nut & 30 & 0.01\euro & 0.30\euro \\
PCB\footnote{Price based on calculation at \url{http://multi-circuit-boards.eu} with the PCB dimensions, double-sided with solder stop.} & 1 & 3.66\euro & 3.66\euro \\ \hline\hline
Total cost& & & 31.28\euro \\
\bottomrule
\caption{Bill of Material for required parts (assuming mass production of 100 units)}
\label{tbl:bom}
\end{longtable}

\begin{longtable}{p{0.5\textwidth}p{0.15\textwidth}p{0.2\textwidth}p{0.1\textwidth}}
\toprule
Part & Amount & Cost per part\footnote{Items supplied via Chinese distributors found on \url{http://alibaba.com} on the 16th of March 2014. The price is based on quantities of one, dollar prices were converted to Euro with 1\euro = 1.3USD } & Sum \\
\midrule
DISPLAY SPI 1.8 INCH & 1 & 4.47\euro & 4.47\euro \\
NRF24L01+ & 1 & 0.92\euro & 0.92\euro \\
MPU6050 GYRO + ACCEL & 1 & 2.76\euro & 2.76\euro \\
HMC-5883L & 1 & 1.84\euro & 1.84\euro \\
MICRO SERVO & 2 & 2.04\euro & 2.04\euro \\
CONN 2.54 FEMALE 1x04 & 1 & 0.04\euro & 0.04\euro \\
CONN 2.54 FEMALE 1x05 & 1 & 0.05\euro & 0.05\euro \\
CONN 2.54 FEMALE 1x08 & 2 & 0.05\euro & 0.10\euro \\
CONN 2.54 FEMALE 2x04 & 1 & 0.05\euro & 0.05\euro \\
\hline \\
Total cost& & & 12.27\euro \\
\bottomrule
\caption{Bill of Material for optional parts (assuming mass production of 100 units)}
\label{tbl:bom_opt}
\end{longtable}
\includepdf[pages={1}]{robot.pdf}
\includepdf[pages={1}, angle=90]{../hardware/robot_schematic.pdf}
